Direct Spectroscopic Detection: A Cheap Method to Detect and Characterize Binary Systems

### 2015-10-09 paper/ms.tex
Updated Section 6.

We use the same set of instruments and settings for all observations throughout the three datasets. We use the CHIRON spectrograph \citep{CHIRON} on the 1.5m telescope at Cerro Tololo Inter-American Observatory for all southern targets. This spectrograph is an $R\equiv \lambda / \Delta \lambda = 80000$ cross-dispersed echelle spectrograph with wavelength coverage from 450 - 850 nm. The data are automatically reduced with a standard CHIRON data reduction pipeline, but the pipeline leaves residuals of strong lines in adjacent orders. We bias-correct, flat-field and extract the spectra with the optimum extraction technique \citep{Horne1986} using standard IRAF\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.} tasks, and use the wavelength calibration from the pipeline reduced spectra.

@@

We now use the direct spectroscopic detection method to measure the temperatures of several known binary systems (Table \ref{tab:known}). We cross-correlate the spectra against the full grid of model spectra enumerated in Section \ref{sec:method}, and find the temperature using Equation \ref{eqn:tmeas}. We then convert the measured temperature to PDFs for the true companion temperature using the MCMC chains developed in Section \ref{subsec:systematics}. For stars with more than one detection, we multiply the PDFs from each detection. Finally, we calculate the final companion temperature and confidence interval from the integral of the PDF:

\begin{equation}
f = \int_{-\infty}^x {pdf(T)dT}
\end{equation}
We use as the central value the value of x such that $f=0.5$ (the median). Likewise, we calculate the $1 \sigma$ lower and upper bounds such that $f = 0.16$ and $f = 0.84$, respectively. The CCFs for the companions that we detect are shown in Figures \ref{fig:detections} and \ref{fig:detections2}. For each star, we show the CCF which has the maximum peak value and annotate the figures with the parameters.

@@

\citet{Zorec2012} find fundamental parameters for HIP 22958, and determine a temperature slightly cooler and luminosity much greater than the spectral type (B6V) would suggest. Because of this, the usual analysis which uses main sequence relationships results in a biased answer. We estimate the companion temperature by assuming that the companion \emph{does} follow the main sequence relationships as described in Section \ref{sec:obs}, but sample the uncertainty distributions given in \citet{Zorec2012} for the temperature and radius of the primary star.

We compare our companion temperature measurements from the direct spectral detection method to the estimates described above in Figure \ref{fig:known}. There is overall excellent agreement between the temperatures, with most points falling within $1 \sigma$ of equality. We test for a bias ($\Delta$) between the measured temperatures ($T_m$) and expected temperatures ($T_e$) with the equations

\begin{eqnarray}
\Delta = \sum_iT_{m,i} - T_{a,i}
\sigma_{\Delta}^2 = \sum_i \sigma_{T_{m,i}}^2 + \sigma_{T_{a,i}}^2
\end{eqnarray}
which results in $\Delta = -580 \pm 890$. Our temperature measurements are consistent with the expected temperatures.

@@

\subsection{Non-detections}
\label{sec:nondetections}
There are many companions in Table \ref{tab:known} that we do not detect. Most of these are single-lined spectroscopic binaries (Table \ref{tab:specdata}), and are likely too cool to detect with our data; very high signal-to-noise spectra with a near-infrared instrument such as IGRINS may uncover them. Several of the remaining un-detected companions have expected temperatures $T  > 6250$ K, the canonical limit at which the convective zone is too small to transfer angular momentum to the stellar wind and spin down the star \citep{Pinsonneault2001}. Since the cross-correlation function gets most of its power from  sharp spectral features, these rapidly rotating companions are difficult to detect (see Figure \ref{fig:sensitivity_2d}).

Finally, HIP 88290 is hot enough and expected to be rotating slowly enough that we should be able to easily detect it. In fact, we would expect to be able to directly see the companion in the spectra (the green lines in Figure \ref{fig:expected}). The fact that we do not see the composite spectrum or see a peak in the cross-correlation function implies that the companion must be rotating with $v\sin{i} > 50 \rm km s^{-1}$, much more quickly that Equation \ref{eqn:gyro} predicts, or that the primary is a giant and therefore much brighter than main-sequence relationships suggest. However this star is in the \cite{David2015} sample and has an effective temperature and mass consistent with main sequence, so we can rule out the latter possibility.

\begin{figure}
        \centering
        \includegraphics[width= 80mm]{HIP_88290_Flux.pdf}
         \caption{Observed (black) and expected (green) spectra for the known binary system HIP 88290. At the expected flux ratio, the spectral lines from the companion should be easily visible. The fact that we do not detect the companion implies that it is either fell outside of the spectrograph slit, or that the companion is a very rapid rotator.}
         \label{fig:expected}
\end{figure}

\section{Discussion and Conclusions}
\label{sec:conclusions}
We have presented and extensively characterized the direct spectral detection method for finding companions to intermediate-mass stars using high resolution cross-dispersed echelle spectroscopy. Using a very large number of synthetic but realistic binary star observations, we have constrained the uncertainty and systematic errors present in determining the companion temperature with the direct spectral detection method. The typical uncertainties are on the order of 200 K across all instruments used in this study, with a systematic offset of similar magnitude (for the optical instruments). We use the synthetic binary star analysis to calibrate the direct spectral detection method for the four instruments used in this study between the temperatures $ 3500 < T < 6500$.

@@

\enddata
%\vspace{-0.8cm}
\tablecomments{The orbital data is from the SB9 database \citep{Pourbaix2004}, and the original references are provided as superscripts after the star names: [1]: \cite{Hill1971}; [2]: \cite{Lloyd1981}; [3]: \cite{Rucinski2005}; [4]: \cite{Abt1978} [5]: \cite{Pourbaix2004}; [6]: \cite{Morrell1992}; [7]: \cite{Abt1990}; [8]: \cite{Fekel1982}; [9]: \cite{Lucy1971}; [10]: \cite{Pogo1928}; [11]: \cite{Duerbeck1975}; [12]: \cite{Abt1965}; [13]: \cite{Scarfe2000}; [14]: \cite{Tomkin1986}; [15]: \cite{Levato1987}; [16]: \cite{Richardson1957}; [17]: \cite{Leone1999}; [18]: \cite{Hube1973};  [19]: \cite{Pearce1936}
}
\end{deluxetable}

@@

\enddata
%\vspace{-0.8cm}
\tablecomments{The imaging data comes from the Washington Double Star Catalog \citep{WDS}, and the most recent measurements are given as superscripts to the star name: [1]: \cite{McAlister1989}; [2]: \cite{Roberts2007}; [3]: \cite{Hipparchos}; [4]: \cite{Mamajek2010}; [5]: \cite{Drummond2014}; [6]: \cite{Shatsky2002}; [7]: \cite{Tokovinin2010}; [8]: \cite{McAlister1987}; [9]: \cite{Horch2010}; [10]: \cite{Horch2008}; [11]: \cite{Horch2001}; [12]: \cite{DeRosa2012}
}
\end{deluxetable}

>>>>

We use the same set of instruments and settings for all observations throughout the three datasets. We use the CHIRON spectrograph \citep{CHIRON} on the 1.5m telescope at Cerro Tololo Inter-American Observatory for all southern targets. This spectrograph is an $R\equiv \lambda / \Delta \lambda = 80000$ cross-dispersed echelle spectrograph with wavelength coverage from 450 - 850 nm, and is fed by a $2.7''$ optical fiber. The data are automatically reduced with a standard CHIRON data reduction pipeline, but the pipeline leaves residuals of strong lines in adjacent orders. We bias-correct, flat-field and extract the spectra with the optimum extraction technique \citep{Horne1986} using standard IRAF\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.} tasks, and use the wavelength calibration from the pipeline reduced spectra.

@@

We now use the direct spectroscopic detection method to measure the temperatures of several known binary systems (Table \ref{tab:known}). We cross-correlate the spectra against the full grid of model spectra enumerated in Section \ref{sec:method}, and find the temperature using Equation \ref{eqn:tmeas}. We then convert the measured temperature to PDFs of the true companion temperature using the MCMC chains developed in Section \ref{subsec:systematics} (see also Figure \ref{fig:error}). For stars with multiple observations, we multiply the PDFs from each detection. Finally, we calculate the companion temperature and confidence interval from the integral of the PDF:

\begin{equation}
f = \int_{-\infty}^x {pdf(T)dT}
\end{equation}
We use as the central value the value of x such that $f=0.5$ (the median). Likewise, we calculate the $1 \sigma$ lower and upper bounds such that $f = 0.16$ and $f = 0.84$, respectively. The CCFs for the companions that we detect are shown in Figures \ref{fig:detections} and \ref{fig:detections2}. For each star, we show the CCF which has the maximum peak value and annotate the figures with the parameters. Most of the CCFs have very strong peaks. The exception is HIP 22958; however, the detection is strengthened by the fact that we observed this star twice and measured a similar temperature both times. The CCFs for HIP 22958 and HIP 24902 demonstrate the adverse effect a large companion rotational velocity has on the detection significance.

@@

\citet{Zorec2012} find fundamental parameters for HIP 22958, and determine a temperature slightly cooler and luminosity much greater than the spectral type (B6V) would suggest. Because of this the usual analysis, which uses main sequence relationships, results in a biased answer. We estimate the companion temperature by assuming that the companion \emph{does} follow the main sequence relationships as described in Section \ref{sec:obs}, but sample the uncertainty distributions given in \citet{Zorec2012} for the temperature and radius of the primary star.

We compare our companion temperature measurements from the direct spectral detection method to the estimates described above in Figure \ref{fig:known}. There is overall excellent agreement between the temperatures, with most points falling within $1 \sigma$ of equality. We test for a bias ($\Delta$) between the measured temperatures ($T_m$) and expected temperatures ($T_a$) with the equations

\begin{eqnarray}
\Delta &= \sum_i(T_{m,i} - T_{a,i}) \\
\sigma_{\Delta}^2 &= \sum_i (\sigma_{T_{m,i}}^2 + \sigma_{T_{a,i}}^2)
\end{eqnarray}
which results in $\Delta = -580 \pm 890$ K. Our temperature measurements are consistent with the expected temperatures.

@@

\subsection{Non-detections}
\label{sec:nondetections}
There are many companions in Table \ref{tab:known} that we do not detect. Most of these are single-lined spectroscopic binaries (Table \ref{tab:specdata}), and are likely too cool to detect with our data; very high signal-to-noise spectra with a near-infrared instrument such as IGRINS may uncover them. Several of the remaining un-detected companions have expected temperatures $T  > 6250$ K, and so are likely to be rapid rotators. Since the cross-correlation function gets most of its power from  sharp spectral features, these rapidly rotating companions are difficult to detect (see Figure \ref{fig:sensitivity_2d}).

Finally, HIP 88290 is hot enough and expected to be rotating slowly enough that we should be able to easily detect it. In fact, we would expect to be able to directly see the companion in the spectra (the green lines in Figure \ref{fig:expected}). The fact that we do not see the composite spectrum or see a peak in the cross-correlation function implies that the companion must be rotating with $v\sin{i} > 50 \rm km s^{-1}$, much more quickly than Equation \ref{eqn:gyro} predicts, that the primary is a giant and therefore much brighter than main-sequence relationships suggest, or that the companion fell outside the spectrograph slit. This star is in the \cite{David2015} sample and has an effective temperature and mass consistent with main sequence, so we can rule out the giant primary possibility. Additionally, the binary separation is $0.47''$ \citep{Tokovinin2015} and the CHIRON spectrograph has a $\sim 2.7''$ diameter fiber; light from the companion is guaranteed to fall on the slit.

\begin{figure}
        \centering
        \includegraphics[width= 80mm]{HIP_88290_Flux.pdf}
         \caption{Observed (black) and expected (green) spectra for the known binary system HIP 88290. At the expected flux ratio, the spectral lines from the companion should be easily visible. }
         \label{fig:expected}
\end{figure}

\section{Discussion and Conclusions}
\label{sec:conclusions}
We have presented and extensively characterized the direct spectral detection method for finding companions to intermediate-mass stars using high resolution cross-dispersed echelle spectroscopy. Using a very large number of synthetic but realistic binary star observations, we have constrained the uncertainty and systematic errors present in determining the companion temperature with the direct spectral detection method. The typical uncertainties are on the order of 200 K across all instruments used in this study, with a systematic offset of similar magnitude (for the optical instruments). We use the synthetic binary star analysis to calibrate the direct spectral detection method for the four instruments used in this study between the temperatures $ 3500 K < T < 6500 K$.

@@

\enddata
%\vspace{-0.8cm}
\tablecomments{Known binary stars with spectroscopic orbit solutions. The orbital data is from the SB9 database \citep{Pourbaix2004}, and the original references are provided as superscripts after the star names: [1]: \cite{Hill1971}; [2]: \cite{Lloyd1981}; [3]: \cite{Rucinski2005}; [4]: \cite{Abt1978} [5]: \cite{Pourbaix2004}; [6]: \cite{Morrell1992}; [7]: \cite{Abt1990}; [8]: \cite{Fekel1982}; [9]: \cite{Lucy1971}; [10]: \cite{Pogo1928}; [11]: \cite{Duerbeck1975}; [12]: \cite{Abt1965}; [13]: \cite{Scarfe2000}; [14]: \cite{Tomkin1986}; [15]: \cite{Levato1987}; [16]: \cite{Richardson1957}; [17]: \cite{Leone1999}; [18]: \cite{Hube1973};  [19]: \cite{Pearce1936}
}
\end{deluxetable}

@@

\enddata
%\vspace{-0.8cm}
\tablecomments{Known binary stars detected through either high contrast imaging or interferometry. The imaging data comes from the Washington Double Star Catalog \citep{WDS}, and the most recent measurements are given as superscripts to the star name: [1]: \cite{McAlister1989}; [2]: \cite{Roberts2007}; [3]: \cite{Hipparchos}; [4]: \cite{Mamajek2010}; [5]: \cite{Drummond2014}; [6]: \cite{Shatsky2002}; [7]: \cite{Tokovinin2010}; [8]: \cite{McAlister1987}; [9]: \cite{Horch2010}; [10]: \cite{Horch2008}; [11]: \cite{Horch2001}; [12]: \cite{DeRosa2012}
}
\end{deluxetable}
